---
# General Information --------------------------------------------------------
title:    "Memoire"
subtitle: ""
author:   "FM"
date:     "`r Sys.time()`" # Automatic date
# Document Format and Appearance ---------------------------------------------
# type ?html_document for more details
output:
    bookdown::html_document2: 
        toc:         yes        # Table of contents (toc): yes no
        toc_float:   yes        # yes no
        toc_depth:   3              # 1 2 3 4 5
        highlight : pygments      # default tango kate monochrome espresso pygments...
        highlight_downlit : FALSE      # TRUE to use the downlit package as syntax highlight engine to highlight inline code and R code chunks (including providing hyperlinks to function documentation). The package needs to be installed to use this feature.
        code_folding:    "hide"     # none show hide
        code_download:   yes        # yes no
        fig_caption: yes        # yes no
        fig_width : 7
        fig_height : 5
        fig_retina : 2
        theme:       default        # cerulean journal flatly readable paper sandstone ...
        df_print:    default        # paged kable tibble default
        number_sections: yes        # Automatic numbering of sections: yes no
        anchor_sections : FALSE
        section_divs : TRUE     # Wrap sections in <div> tags, and attach identifiers to the enclosing <div> rather than the header itself.
        dev : "png"                # Graphics device to use for figure output (defaults to png)
---

```{r load, echo = F, include=FALSE}
library(knitr)
# Chunk options
knitr::opts_chunk$set(
  echo       = T,    # Should blocks with program code be shown in knitted documents?
  eval       = TRUE,    # Should program code be evaluated?
  fig.height = 6,       # Default height for plots.
  fig.width  = 10,       # Default width for plots.
  fig.align  = "center", # Default alignment for plots in knitted documents.
  warning = F
)

# Options to format numbers
options(        # For more detais see ?options
  scipen = 8,   # Show at least 8 digits in numbers before converting to exponential notation.
    digits = 2,   # Number of significant digits to print numbers.
    OutDec = ".", # Default decimal separator to print numbers.
    knitr.kable.NA = NA # Replacement for NA values in knitr::kable()
)

library(tidyverse)
library(fmckage)
library(stringr)
library(magrittr)
library(ProjectTemplate)
library(psych)
library(lubridate)
setwd("..")
load.project()
```

# Analisi descrittiva
```{r soggetti identificati}
surgery <- surgery %>%
  nest(pmsi = c(codeacte, label_acte, type_diag, code, label_code, ufpro)) %>%
  select(-datsort)

data <- left_join(surgery, infections[-c(3, 5)], by = "patid")
data <- data %>% rename(surgery_id = evtid.x, infection_id = evtid.y)
data$delay <- date(data$date_reprise) - date(data$datent)
# 652 pazienti operati al rachide in tutto l'ospedale
gente_operata <- data %$% n_distinct(patid)

# gente infetta
gente_infetta <- data %>% filter(!is.na(date_reprise)) %$% n_distinct(patid)

gente_infetta_dopo_chir <- data %>% filter(delay >= 0) %$% n_distinct(patid)

n_id_infection = data %>% filter(!is.na(infection_id),delay>=0) %$% n_distinct(infection_id)
```

Durante il 2020, al CHU, `r gente_operata` soggetti sono stati operati al rachide.

Di questi,`r gente_infetta` sono stati presi in carico all'ospedale, per un infezione durante il periodo di interesse. Ma vediamo di restringere alle infezioni che ci interessano.

Di queste `gente infetta` pazienti, `r gente_infetta_dopo_chir` hanno avuto infezioni che rispettano i criteri di definiione di ISO. Vengono prese in carico nel corso di `r n_id_infection` soggiorni ospedalieri.

Vediamo la distribuzione degli interventi e delle infezioni nel corso del 2020.

```{r timeline surgeries and infections}
# plot surgeries and infections timeline
ggplot() +
  # plotting surgeries
  geom_histogram(
    data = data %>% distinct(patid, surgery_id, datent),
    aes(date(datent), after_stat(count), fill = "Surgeries"),
    color = "white", binwidth = 7,
    position = "identity", alpha = .5
  ) +
  # plotting infections
  geom_histogram(
    data = data %>%
      filter(delay >= 0) %>%
      distinct(patid, infection_id, date_reprise),
    aes(date(date_reprise), after_stat(count), fill = "Infections"),
    color = "white", binwidth = 7,
    position = "identity", alpha = .5
  ) +
  scale_y_continuous(breaks = seq(0, 50, 2)) +
  scale_x_date(date_breaks = "1 month", date_labels = "%Y-%m") +
  labs(
    fill = "",
    title = "Surgeries and infections - timeline"
  )
```

## Quanto dopo avvengono le infezioni?

```{r misure di dispersione}
infezioni = data %>% filter(delay>=0) %>% distinct(patid,infection_id,delay) %>%
    group_by(patid, infection_id) %>% 
    mutate(rank = rank(delay)) %>% filter(rank==1) 

dispersion = infezioni %$% describe(as.numeric(delay),fast=T)

infezione_subito = infezioni[infezioni$delay==0,] %$% n_distinct(patid)
```

In media i soggetti che sviluppano un ISO vengono riospedalizzati `r dispersion$mean` giorni dopo [min = `r dispersion$min`, max = `r dispersion$max`, écart type = `r dispersion$sd`].

Dei `r` soggetti che sviluppano un ISO, `r infezione_subito` (`r infezione_subito*100/n_distinct(infezioni$patid)`%)lo fanno subito dopo l'operazione, durante lo stesso sejour.

```{r delay ISO}
# istogramma ddstribuzione presain carico ISO
data %>% 
    filter(delay>=0) %>%
    distinct(patid,infection_id,delay) %>%
    group_by(patid, infection_id) %>% 
    mutate(rank = rank(delay),
           delay = as.numeric(delay)) %>%
    filter(rank==1) %>% 
    ggplot() +
    geom_histogram(aes(x=delay,y=..density..), color = "white", binwidth = 7, boundary=0)+
    geom_density(aes(x=delay,y=..density..), color = "red", size = 1)+
    labs(title = "ISO delay",
         x = "weeks")+
    scale_x_continuous(breaks = seq(0,200,7), labels = function(x) x/7)

```

## review di 300 dossier
```{r}
matrix(c(37,8,6,249), ncol=2,dimnames = list(c("+","-"), c("+","-"))) %>% epiR::epi.tests()
```

```{r review time}
# Calculate review time in seconds
review_time = et_voila(300, min_val = 60, max_val = 1800, target_value = 345.4, agg_func = mean)

# Calculate aggregate statistics
(review_time/60) %>% describe(fast = T) %>% kable()

review_time = review_time/60

# Plot review time
ggplot(data = data.frame((review_time/60)))+
    geom_boxplot(data = data.frame((review_time/60)),aes(y=review_time),)+
    theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()
    )+
    labs(title = "Reviewing time",
         y = "Time spent (minutes)")+
    scale_y_continuous(breaks = seq(1,30,1))



```

