---
# General Information --------------------------------------------------------
title:    "Thèse - Rachel Frébourg"
subtitle: "Comparaison avant-après niveau interdisciplinarité chez les pro de santé"
author:   "Francesco MONTI"
date:     "2023-06-07 15:54:40" # Automatic date
# Document Format and Appearance ---------------------------------------------
# type ?html_document for more details
output:
    bookdown::html_document2: 
        toc:         yes        # Table of contents (toc): yes no
        toc_float:   yes        # yes no
        toc_depth:   5              # 1 2 3 4 5
        highlight : pygments      # default tango kate monochrome espresso pygments...
        highlight_downlit : FALSE      # TRUE to use the downlit package as syntax highlight engine to highlight 
                                       # inline code and R code chunks (including providing hyperlinks to function 
                                       # documentation). The package needs to be installed to use this feature.
        code_folding:    "hide"     # none show hide
        code_download:   yes        # yes no
        fig_caption: yes        # yes no
        fig_width : 7
        fig_height : 5
        fig_retina : 2
        theme:       default        # cerulean journal flatly readable paper sandstone ...
        df_print:    default        # paged kable tibble default
        number_sections: yes        # Automatic numbering of sections: yes no
        anchor_sections : FALSE
        section_divs : TRUE     # Wrap sections in <div> tags, and attach identifiers to the enclosing <div> 
                                # rather than the header itself.
        dev : "png"                # Graphics device to use for figure output (defaults to png)
        self_contained : TRUE
        extra_dependencies : NULL    # Extra dependencies as a list of the html_dependency class objects typically 
                                     # generated by htmltools:htmlDependency().
        css : style.css     # CSS and/or Sass files to include. Files with an extension of .sass or .scss are compiled to 
                       # CSS via sass::sass(). Also, if theme is a bslib::bs_theme() object, Sass code may reference 
                       # the relevant Bootstrap Sass variables, functions, mixins, etc.
        includes : NULL        # Named list of additional content to include within the document 
                                # (typically created using the includes function)
        keep_md : FALSE        # Keep the markdown file generated by knitting.
        lib_dir : NULL             #    Directory to copy dependent HTML libraries (e.g. jquery, bootstrap, etc.) 
                                   # into. By default this will be the name of the document with _files appended to it.
        md_extensions : NULL       # Markdown extensions to be added or removed from the default definition of 
                                    # R Markdown. See the rmarkdown_format for additional details.
        pandoc_args : NULL      # Additional command line options to pass to pandoc
        template : "default"       # Pandoc template to use for rendering. Pass "default" to use the rmarkdown 
                                    # package default template; pass NULL to use pandoc's built-in template; 
                                    # pass a path to use a custom template that you've created. 
                                    # Note that if you don't use the "default" template then some features of 
                                    # html_document won't be available (see the Templates section below for more details).
        math_method : "default"     # Math rendering engine to use. This will define the math method to use with Pandoc.
                                    # It can be a string for the engine, one of "mathjax", "mathml", "webtex", "katex",
                                    # "gladtex", or "r-katex" or "default" for mathjax.
                                # It can be a list of
                                # - engine: one of "mathjax", "mathml", "webtex", "katex", or "gladtex". 
                                # - url: A specific url to use with mathjax, katex or webtex. 
                            # Note that for engine = "mathjax", url = "local" will use a local version of 
                            # MathJax (which is copied into the output directory).
                            # For example,
                            # output:
                            #    html_document:
                            #        math_method:
                            #        engine: katex
                            #        url: https://cdn.jsdelivr.net/npm/katex@0.11.1/dist
                            #        See Pandoc's Manual about Math in HTML for the details about Pandoc 
                            #        supported methods.

                        # Using math_method = "r-katex" will opt-in server side rendering using KaTeX thanks to 
                        # katex R package. 
                        # This is useful compared to math_method = "katex" to have no JS dependency, 
                        # only a CSS dependency for styling equation.
        mathjax : "default"        # Include mathjax. The "default" option uses an https URL from a MathJax CDN. 
                                    # The "local" option uses a local version of MathJax (which is copied into 
                                    # the output directory). You can pass an alternate URL or pass NULL 
                                    # to exclude MathJax entirely.
---



```{r chunk options, echo=F}
# Chunk options
knitr::opts_chunk$set(
  echo       = T,    # Should blocks with program code be shown in knitted documents?
  eval       = TRUE,    # Should program code be evaluated?
  fig.height = 6,       # Default height for plots.
  fig.width  = 10,       # Default width for plots.
  fig.align  = "center", # Default alignment for plots in knitted documents.
  warning = F
)

```

```{r libraries and data loading, include = F}

# LIBRARIES ----------------------------------------------------------------------------------
library(conflicted)    # Get a warning/error if several functions with the same name exist.
library(magrittr)      # Operator %>% and additional pipe-friendly functions.
library(tidyverse)     # The main "tidyverse" packages.
library(openxlsx)      # Write data to Excel files.
library(stringr)
library(knitr)
library(kableExtra)
library(pander)
library(DT)
library(janitor)

conflicted::conflicts_prefer(dplyr::filter)

# DATA ------------------------------------------------------------------------------------------
post = read.xlsx("Copie de resultats Q2 copie.xlsx", sheet = 1)
pre = read.xlsx("Copie de resultats Q1 copie.xlsx", sheet = 1)
```

```{r data management, include = F}
# Colnames() to lowercase
colnames(pre) = tolower(colnames(pre))
colnames(post) = tolower(colnames(post))

# MERGING PRE E POST ----------------------------------------------------------
pre = pre %>% 
    select(-contains(c("id","submitdate","lastpage","startlanguage"))) %>% 
    rename("filiere"=b1, "id" = a1) %>% 
    mutate(test = "pre")

post = post %>%
    filter(!is.na(submitdate)) %>% 
    select(-contains(c("id","submitdate","lastpage","startlanguage","x26","x27"))) %>% 
    rename("id" = a1) %>% 
    left_join(pre[,c("id","filiere")], by = "id") %>% 
    relocate(filiere, .after = "id") %>% 
    mutate(test = "post")

# Merging
data = rbind(pre,post) %>% tibble

# DATA MANAGEMENT merged DF ----------------------------------------------------
data = data %>% 
    mutate(across(s1q1:s7q2, as.numeric)) %>% 
    distinct() %>% 
    relocate(test, .after = "id") %>%
    mutate(filiere = case_when(filiere == "étudiant(e) en maïeutique" ~ "maieutique",
                               filiere == "étudiant(e) en orthophonie" ~ "orthophonie",
                               filiere == "étudiant(e) en pharmacie" ~ "pharmacie",
                               filiere == "étudiant(e) IDE" ~ "IDE",
                               filiere == "étudiant(e) MKE" ~ "MKE",
                               filiere == "interne de médecine générale" ~ "medecine generale",
                               TRUE ~ NA_character_
                               )
           ) %>% 
    filter(!is.na(id))


# Computing missing values per row
data = data %>% filter(!is.na(id)) %>% 
    mutate(na_counts = rowSums(across(s1q1:s7q2, is.na)))

# Removing totally incomplete questionnaires
data = data %>% filter(na_counts<20)



```

# Exploratory analysis

## Nb et repartition sujets participants

On a un total de **379** sujets si on regarde les idenfiants anonymes uniques, repartis de la façon suivante:

```{r }
invisible(n_distinct(data$id))

data %>% distinct(id,filiere) %$% 
    table(filiere) %>% 
    as.data.frame() %>% 
    mutate("%" = round(Freq/sum(Freq)*100, 2)) %>% 
    adorn_totals() %>% 
    pander(caption = "**Repartition sujets par filière**")
```

<br>

<br>

## Missing values

Voyons sur tout l'ensemble de tests, combien de données manquantes on a. Ce sont des "vraies" données manquantes **OU** des "NSP" qui n'était pas pertinents pour la question et vraisemblablement pris pour des "Ne sait pas".

```{r missing values}


table(data$na_counts) %>% 
    as.data.frame() %>% 
    mutate("%" = round(Freq/sum(Freq)*100, 2)) %>% 
    rename("Nb missing values" = Var1) %>% 
    pander(caption = "**Nb de valeurs manquantes par questionnaire**")

```

On voit que les tests sont bien complets globalement sauf 16 questionnaires qui sont **complétement** vides. Ils sont des doublons dus à des enregistrements vides.

<br>

<br>

### Missing values par filière

On peut voir que la repartition des valeurs manquantes par filière est en ligne avec la repartition des sujets par filière. Donc il n'y a pas une filière qui a laissé le questionnaire "plus vide" que les autres.

```{r}
data %>%
  group_by(filiere) %>%
  summarise(Freq = sum(na_counts)) %>%
  mutate("%" = round(Freq / sum(Freq) * 100, 2)) %>%
    adorn_totals() %>% 
  pander(caption = "**Valeurs manquantes par filière**", align = "center")


```

**!!!!!!** Quelque chose ne va pas, on a 379 sujets et 800+ questionnaires (on devrait en avoir **au max** 379\*2). Dans le tableau suivant on voit qu'on a 199 sujets avec **au moins** 3 entrées (et j'ai déjà exclu les 16 dont je parlais avant).

```{r message = F}
x <- data %>%
  group_by(id, filiere) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  filter(n > 2) %>%
  pull(id) 

data %>%
    filter(id %in% x) %>%
    arrange(id, filiere, test) %>%
    datatable(options = list(pageLength = 10,
                             #dom = 't',
                             scrollX = TRUE,
                             width = "auto",
                             scrollX = T
                             )
              ) 
```

Les entrées supplementaires ne sont pas des "doublons", les valeurs ne sont pas les mêmes. Je vais donc prendre la moyenne de ces valeurs. Ci dessous le résultat, on est à 766 lignes alor que notre max theorique est 758. Il y a forcement des gens qui se sont enregistrés avec plus qu'un ID (j'ai renommé la variable "A1").   
Ca ne m'inquiète pas pour les tests appariés, je ne pense pas que des gens ont 2*2 tests appariés mais je controllerai.

```{r}
data %>% 
    group_by(id,test,filiere) %>% 
    summarise(across(s1q1:s7q2,mean,na.rm=T)) %>% 
    mutate(na_counts = rowSums(across(s1q1:s7q2, is.na))) %>% 
    datatable(options = list(pageLength = 10,
                             #dom = 't',
                             scrollX = TRUE,
                             width = "auto",
                             scrollX = T
                             )
              )
```

<br>

<br>

# Tests appariés

```{r Cas appariés par filière}
# DF with id-filiere-pre-post pour les tests stat
test_data <- data %>%
  group_by(id, filiere, test) %>%
  summarise(mean_score = mean(c_across(s1q1:s7q2), na.rm = T)) %>%
  pivot_wider(names_from = "test", values_from = "mean_score") %>%
    filter(!is.na(post),!is.na(pre)) %>% distinct

# paired cases by filiere
test_data %$% 
    table(filiere) %>% 
    as.data.frame() %>% 
    mutate("%" = round(Freq/sum(Freq)*100, 2)) %>% 
    adorn_totals() %>% 
    pander(caption = "Cas appariés par filière")
```

Pour ces pourcentages les denominateur est "total de tests appariés" et non le "nombre total de tests. Donc, les IDE, representent 38.14% des tests appariés disponibles.

Mais, il y a carrement plus d'IDE que les autres filières! Oui, bien sur. Voyons

# Tests stat

```{r analysis}
# PRE vs POST, toutes filieres confondues
t.test(test_data$post,test_data$pre, paired=T)

# PRE vs POST, filiere par filiere 
t.test(test_data$post[test_data$filiere=="IDE"],test_data$pre[test_data$filiere=="IDE"], paired=T)

```
